---
title: "02_survival-analysis"
output: html_document
date: "2026-01-30"
---

## Survival Analysis

```{r setup, echo=FALSE}

library(tidyverse)
library(survival)
library(data.table)


ntb <- fread("../data/in-progress/ntb_v5.csv")
tb <- fread("../data/in-progress/tb_v6.csv")

# Important functions

remove_blank <- function(x){
  x <- x |>
    mutate(across(everything(), ~if_else(.x == "", NA, .x)))
  return(x)
}

cln_dupl <- function(x){
  x[duplicated(x)] <- NA
  return(x)
}
```

Now let's try to fit a Kaplan-Meier curve to the people with tuberculosis mentioned in their record.

Start by finding the first date that is associated with a mention of TB

```{r}

tb$recidnum <- as.character(tb$recidnum)

tb_date_date <- remove_blank(tb) |>
  select(recidnum,
    #a_clm01_date = appdt01,
    #a_clm02_date = appdt02,
    #a_clm03_date = appdt03,
    #a_clm04_date = appdt04,
    #a_clm05_date = appdt05,
    #a_clm07_date = appdt07,
    #a_clm08_date = appdt08,
    #a_clst01_date = appdt01,
    #a_clst02_date = appdt02,
    #a_clst07_date = appdt07,
    #afdesc01_date = infodt01,
    #afdesc02_date = infodt02,
    #afdesc07_date = infodt07,
    dschrn01_date = dschdt01,
dschrn02_date = dschdt02,
dschrn03_date = dschdt03,
dschrn04_date = dschdt04,
dschrn05_date = dschdt05,
dschrn07_date = dschdt07,
dschrn08_date = dschdt08,
dschrn09_date = dschdt09,
dschrn10_date = dschdt10,
dschrn11_date = dschdt11,
dschrn12_date = dschdt12,
dschrn13_date = dschdt13,
dschrn14_date = dschdt14,
dschrn16_date = dschdt16,
dschrn19_date = dschdt19,
dschrn20_date = dschdt20,
ilwdsc01_date = admtdt01,
ilwdsc02_date = admtdt02,
ilwdsc03_date = admtdt03,
ilwdsc04_date = admtdt04,
ilwdsc05_date = admtdt05,
ilwdsc06_date = admtdt06,
ilwdsc07_date = admtdt07,
ilwdsc08_date = admtdt08,
ilwdsc09_date = admtdt09,
ilwdsc10_date = admtdt10,
ilwdsc11_date = admtdt11,
ilwdsc12_date = admtdt12,
ilwdsc13_date = admtdt13,
ilwdsc14_date = admtdt14,
ilwdsc15_date = admtdt15,
ilwdsc16_date = admtdt16,
ilwdsc17_date = admtdt17,
ilwdsc18_date = admtdt18,
ilwdsc19_date = admtdt19,
ilwdsc20_date = admtdt20,
othrrn01_date = othrdt01,
othrrn02_date = othrdt02,
unftc1_1_date = dchdt1_1,
unftc1_2_date = dchdt1_2,
unftc1_3_date = dchdt1_3,
unftc2_1_date = dchdt2_1
  ) |>
  mutate(across(-recidnum, as.character))


tb_date_event <- remove_blank(tb) |>
  select(recidnum,
dschrn01,
dschrn02,
dschrn03,
dschrn04,
dschrn05,
dschrn07,
dschrn08,
dschrn09,
dschrn10,
dschrn11,
dschrn12,
dschrn13,
dschrn14,
dschrn16,
dschrn19,
dschrn20,

ilwdsc01,
ilwdsc02,
ilwdsc03,
ilwdsc04,
ilwdsc05,
ilwdsc06,
ilwdsc07,
ilwdsc08,
ilwdsc09,
ilwdsc10,
ilwdsc11,
ilwdsc12,
ilwdsc13,
ilwdsc14,
ilwdsc15,
ilwdsc16,
ilwdsc17,
ilwdsc18,
ilwdsc19,
ilwdsc20,

othrrn01,
othrrn02,

unftc1_1,
unftc1_2,
unftc1_3,
unftc2_1
  )



# Lengthen the dataset and identify pairs

event_long <- tb_date_event |>
  pivot_longer(
    cols = -recidnum,
    names_to = "slot",
    values_to = "event"
  )

date_long <- tb_date_date |>
  pivot_longer(
    cols = -recidnum,
    names_to = "slot",
    values_to = "date"
  ) |>
  mutate(slot = str_remove(slot, "_date$")) 

tb_long <- event_long |>
  inner_join(date_long, by = c("recidnum", "slot"))

tb_dates <- tb_long |>
  filter(
    !is.na(event),
    str_detect(event, regex("TUBER|CONSUMP|PHTH", ignore_case = TRUE))
  ) |>
  group_by(recidnum) |>
  summarise(
    tb_dates = list(date),
    .groups = "drop"
  )


###### And then join everything back to tb

tb_min_dates <- tb_dates |>
  mutate(
    tb_dates_clean = map(tb_dates, ~ .x[!is.na(.x) & .x != ""]),
    # Replace trailing "--" with "15"
    tb_dates_clean = map(tb_dates_clean, ~ str_replace(.x, "--$", "15")),
    # Keep only valid 8-digit dates
    tb_dates_clean = map(tb_dates_clean, ~ .x[str_detect(.x, "^\\d{8}$")]),
    # Get earliest date
    tb_first_date = map_chr(tb_dates_clean, ~ if(length(.x) == 0) NA_character_ else min(.x))
  ) |>
  select(recidnum, tb_first_date)

tb_min_dates$recidnum <- as.character(tb_min_dates$recidnum)

first_tb_full <- tb |>
  left_join(tb_min_dates, by = "recidnum") |>
  select(recidnum,
         BYR,
         DYR,
         tb_first_date)

```

Now let's get a data frame with relevant date by recruit

```{r}



first_tb_full <- first_tb_full |>
  mutate(recidnum = as.character(recidnum))

first_tb_full <- first_tb_full |>
  mutate(
    tb_first_date = substr(tb_first_date, 1, 4)
  )

first_tb_full <- remove_blank(first_tb_full) |>
  drop_na() |>
  mutate(
    DYR = as.numeric(DYR),
    tb_first_date = as.numeric(tb_first_date)
  )

glimpse(first_tb_full)


```

And plot!

```{r}

library(survival)
library(dplyr)


# Prepare dataset
surv_tb <- first_tb_full |>
  mutate(
    age_tb = tb_first_date - BYR,
    age_death = DYR - BYR
  )

# Right-censored survival object (referencing the columns inside the dataframe)
surv_obj <- Surv(
  time = ifelse(!is.na(surv_tb$age_tb), surv_tb$age_tb, surv_tb$age_death),
  event = ifelse(!is.na(surv_tb$age_tb), 1, 0)
)

# Fit Kaplan-Meier
km_fit <- survfit(surv_obj ~ 1)

# Base R plot
plot(km_fit,
     xlab = "Age (years)",
     ylab = "Survival with TB",
     main = "Kaplan-Meier Curve",
     col = "blue",
     lwd = 2)

# Optional: add confidence intervals
lines(km_fit$time, km_fit$upper, lty = 2, col = "red")
lines(km_fit$time, km_fit$lower, lty = 2, col = "red")

print(km_fit)


# Calculate time since TB mention and censoring
surv_tb <- first_tb_full %>%
  filter(!is.na(tb_first_date)) %>%  # only people who had TB
  mutate(
    start_age = tb_first_date - BYR,    # age at first TB
    end_age = DYR - BYR,                # age at death
    time_since_tb = end_age - start_age, # duration after TB
    event = 1                           # everyone eventually dies, so event = 1
  ) %>%
  filter(time_since_tb >= 0)            # remove negative intervals if any

surv_obj <- Surv(time = surv_tb$time_since_tb, event = surv_tb$event)

km_fit <- survfit(surv_obj ~ 1)

plot(km_fit,
     xlab = "Years since first TB mention",
     ylab = "Survival probability",
     main = "Kaplan-Meier Curve: Survival After TB",
     col = "blue",
     lwd = 2)

# Optional: add confidence intervals
lines(km_fit$time, km_fit$upper, lty = 2, col = "red")
lines(km_fit$time, km_fit$lower, lty = 2, col = "red")

print(km_fit)

hist(surv_tb$time_since_tb)
hist(surv_tb$end_age)
hist(surv_tb$tb_first_date)


```

To be able to compare TB to non-TB soldiers, we need to evaluate survival from a common event. Let's do first enlistment, which will control (slightly) for age.

```{r lst_km}




```
