---
title: "02_survival-analysis"
output: html_document
date: "2026-01-30"
---

## Survival Analysis

```{r setup, echo=FALSE}

library(tidyverse)
library(survival)
library(survminer)
library(data.table)
library(MatchIt)
library(knitr)

ntb <- fread("../data/in-progress/ntb_v6.csv")
tb <- fread("../data/in-progress/tb_v7.csv")
all <- rbind(tb, ntb)

# Important functions

remove_blank <- function(x){
  x <- x |>
    mutate(across(everything(), ~if_else(.x == "", NA, .x)))
  return(x)
}

cln_dupl <- function(x){
  x[duplicated(x)] <- NA
  return(x)
}
```

## Summary Statistics

```{r}




```

Now let's try to fit a Kaplan-Meier curve to the people with tuberculosis mentioned in their record.

Start by finding the first date that is associated with a mention of TB

```{r}

all$recidnum <- as.character(all$recidnum)

tb_date_date <- remove_blank(all) |>
  select(recidnum,
    #a_clm01_date = appdt01,
    #a_clm02_date = appdt02,
    #a_clm03_date = appdt03,
    #a_clm04_date = appdt04,
    #a_clm05_date = appdt05,
    #a_clm07_date = appdt07,
    #a_clm08_date = appdt08,
    #a_clst01_date = appdt01,
    #a_clst02_date = appdt02,
    #a_clst07_date = appdt07,
    #afdesc01_date = infodt01,
    #afdesc02_date = infodt02,
    #afdesc07_date = infodt07,
    dschrn01_date = dschdt01,
dschrn02_date = dschdt02,
dschrn03_date = dschdt03,
dschrn04_date = dschdt04,
dschrn05_date = dschdt05,
dschrn07_date = dschdt07,
dschrn08_date = dschdt08,
dschrn09_date = dschdt09,
dschrn10_date = dschdt10,
dschrn11_date = dschdt11,
dschrn12_date = dschdt12,
dschrn13_date = dschdt13,
dschrn14_date = dschdt14,
dschrn16_date = dschdt16,
dschrn19_date = dschdt19,
dschrn20_date = dschdt20,
ilwdsc01_date = admtdt01,
ilwdsc02_date = admtdt02,
ilwdsc03_date = admtdt03,
ilwdsc04_date = admtdt04,
ilwdsc05_date = admtdt05,
ilwdsc06_date = admtdt06,
ilwdsc07_date = admtdt07,
ilwdsc08_date = admtdt08,
ilwdsc09_date = admtdt09,
ilwdsc10_date = admtdt10,
ilwdsc11_date = admtdt11,
ilwdsc12_date = admtdt12,
ilwdsc13_date = admtdt13,
ilwdsc14_date = admtdt14,
ilwdsc15_date = admtdt15,
ilwdsc16_date = admtdt16,
ilwdsc17_date = admtdt17,
ilwdsc18_date = admtdt18,
ilwdsc19_date = admtdt19,
ilwdsc20_date = admtdt20,
othrrn01_date = othrdt01,
othrrn02_date = othrdt02,
unftc1_1_date = dchdt1_1,
unftc1_2_date = dchdt1_2,
unftc1_3_date = dchdt1_3,
unftc2_1_date = dchdt2_1
  ) |>
  mutate(across(-recidnum, as.character))


tb_date_event <- remove_blank(all) |>
  select(recidnum,
dschrn01,
dschrn02,
dschrn03,
dschrn04,
dschrn05,
dschrn07,
dschrn08,
dschrn09,
dschrn10,
dschrn11,
dschrn12,
dschrn13,
dschrn14,
dschrn16,
dschrn19,
dschrn20,

ilwdsc01,
ilwdsc02,
ilwdsc03,
ilwdsc04,
ilwdsc05,
ilwdsc06,
ilwdsc07,
ilwdsc08,
ilwdsc09,
ilwdsc10,
ilwdsc11,
ilwdsc12,
ilwdsc13,
ilwdsc14,
ilwdsc15,
ilwdsc16,
ilwdsc17,
ilwdsc18,
ilwdsc19,
ilwdsc20,

othrrn01,
othrrn02,

unftc1_1,
unftc1_2,
unftc1_3,
unftc2_1
  )



# Lengthen the dataset and identify pairs

event_long <- tb_date_event |>
  pivot_longer(
    cols = -recidnum,
    names_to = "slot",
    values_to = "event"
  )

date_long <- tb_date_date |>
  pivot_longer(
    cols = -recidnum,
    names_to = "slot",
    values_to = "date"
  ) |>
  mutate(slot = str_remove(slot, "_date$")) 

tb_long <- event_long |>
  inner_join(date_long, by = c("recidnum", "slot"), relationship = "many-to-many")

tb_dates <- tb_long |>
  filter(
    !is.na(event),
    str_detect(event, regex("TUBER|CONSUMP|PHTH", ignore_case = TRUE))
  ) |>
  group_by(recidnum) |>
  summarise(
    tb_dates = list(date),
    .groups = "drop"
  )


###### And then join everything back to tb

tb_min_dates <- tb_dates |>
  mutate(
    tb_dates_clean = map(tb_dates, ~ .x[!is.na(.x) & .x != ""]),
    # Replace trailing "--" with "15"
    tb_dates_clean = map(tb_dates_clean, ~ str_replace(.x, "--$", "15")),
    # Keep only valid 8-digit dates
    tb_dates_clean = map(tb_dates_clean, ~ .x[str_detect(.x, "^\\d{8}$")]),
    # Get earliest date
    tb_first_date = map_chr(tb_dates_clean, ~ if(length(.x) == 0) NA_character_ else min(.x))
  ) |>
  select(recidnum, tb_first_date)

tb_min_dates$recidnum <- as.character(tb_min_dates$recidnum)

first_tb_full <- all |>
  left_join(tb_min_dates, by = "recidnum") |>
  select(recidnum,
         BYR,
         DYR,
         tb_first_date,
         TB_yn,
         
         # For later matching
         initstat, initreg, gen_birthplace_icpsr1, wnd_yn
         )

```

Now let's get a data frame with relevant date by recruit

```{r df_tb}



first_tb_full <- first_tb_full |>
  mutate(recidnum = as.character(recidnum))

first_tb_full <- first_tb_full |>
  mutate(
    tb_first_date = substr(tb_first_date, 1, 4)
  )

first_tb_full <- remove_blank(first_tb_full) |>
  mutate(
    DYR = as.numeric(DYR),
    tb_first_date = as.numeric(tb_first_date)
  )

glimpse(first_tb_full)

# Prepare dataset for survival curve
surv_tb <- first_tb_full |>
  mutate(
    age_tb = tb_first_date - BYR,
    age_death = DYR - BYR
  )



```

And plot!

```{r km_tb}

library(survival)
library(dplyr)


# Right-censored survival object (referencing the columns inside the dataframe)
surv_obj <- Surv(
  time = ifelse(!is.na(surv_tb$age_tb), surv_tb$age_tb, surv_tb$age_death),
  event = ifelse(!is.na(surv_tb$age_tb), 1, 0)
)

# Fit Kaplan-Meier
km_fit <- survfit(surv_obj ~ 1)

# Base R plot
plot(km_fit,
     xlab = "Age (years)",
     ylab = "Survival with TB",
     main = "Kaplan-Meier Curve",
     col = "blue",
     lwd = 2)

# Optional: add confidence intervals
lines(km_fit$time, km_fit$upper, lty = 2, col = "red")
lines(km_fit$time, km_fit$lower, lty = 2, col = "red")

print(km_fit)


# Calculate time since TB mention and censoring
surv_tb_2 <- first_tb_full |>
  filter(!is.na(tb_first_date)) |>  # only people who had TB
  mutate(
    start_age = tb_first_date - BYR,    # age at first TB
    end_age = DYR - BYR,                # age at death
    time_since_tb = end_age - start_age, # duration after TB
    event = 1                           # everyone eventually dies, so event = 1
  ) |>
  filter(time_since_tb >= 0)            # remove negative intervals if any

surv_obj <- Surv(time = surv_tb_2$time_since_tb, event = surv_tb_2$event)

km_fit <- survfit(surv_obj ~ 1)

plot(km_fit,
     xlab = "Years since first TB mention",
     ylab = "Survival probability",
     main = "Kaplan-Meier Curve: Survival After TB",
     col = "blue",
     lwd = 2)

# Optional: add confidence intervals
lines(km_fit$time, km_fit$upper, lty = 2, col = "red")
lines(km_fit$time, km_fit$lower, lty = 2, col = "red")

print(km_fit)

hist(surv_tb$time_since_tb)
hist(surv_tb$end_age)
hist(surv_tb$tb_first_date)


```

To be able to compare TB to non-TB soldiers, we need to evaluate survival from a common event. Let's do first enlistment, which will control (slightly) for age.

```{r lst_km}


lst <- all |>
  select(recidnum,
    lstdt1_1, mildfdt1, lstdt1_2, mildfdt2,
    lstdt1_3, mildfdt3, lstdt2_1, mildfdt4,
    lstdt2_2, mildfdt5, lstdt2_3, mildfdt6,
    TB_yn, DYR
  )

lst <- as.data.frame(t(apply(remove_blank(lst), 1, cln_dupl)))



# Turn end "--" into 0 and delete every cell that still has "--" to get rid of dates that are too arbitrary (month or year is unknown)

lst <- lst |>
  mutate(
    across(-c(recidnum, TB_yn), ~gsub("--$", "00", .)),
    across(-c(recidnum, TB_yn), ~if_else(grepl("--", .x), NA, .x)),
    across(-c(recidnum, TB_yn), ~ifelse(!is.na(.x), substr(.x, 1, 4), .x))
  ) |>
  filter(if_any(-c(recidnum, TB_yn), ~ !is.na(.x))) |>
  rowwise() |>
  mutate(
    first_lst = min(c_across(-c(recidnum,TB_yn)), na.rm = TRUE)
  ) |>
  ungroup()


# And new lets see the Kaplan-Meier curve for TB versus non-TB

compare_df <- lst |>
  mutate(
    first_lst = as.numeric(first_lst),
    DYR = as.numeric(DYR),
    surv_time = DYR - first_lst,
    event = 1
  ) |>
  filter(surv_time >= 0) a

compare_df$TB_yn <- factor(compare_df$TB_yn, 
                           levels = c(0, 1), 
                           labels = c("Non-TB", "TB"))


km_comparison <- survfit(Surv(surv_time, event) ~ TB_yn, data = compare_df)

print(km_comparison)

p <- ggsurvplot(
  km_comparison, 
  data = compare_df,
  conf.int = TRUE,
  risk.table = FALSE,
  pval = TRUE,
  palette = c("#E7B800", "#2E9FDF"),
  xlab = "Years after First Enlistment",
  ggtheme = theme_minimal()
)

p$plot <- p$plot + labs(title = "Survival: TB vs. Non-TB Cohorts")

p



```

As you can see, the tb vs. non-tb survival curves are significantly different (P \< 0.0001 for log rank test).

The log rank test requires that: - the censoring patterns are the same - the hazard functions for the two groups are proportional

We want to see if their is any causal relationship between the diagnosis of tb. To do this, we will match TB individuals to non-TB individuals with as many of the same characteristics as possible. The characteristics we will try to match are:

-   Birth Year

-   US State of enlistment\

```{r matching 1}

# Characteristics

# Dataframe for matching from the previous time_since_tb dataframe

match_full <- surv_tb |>
  select(
    recidnum,
    tb_first_date, age_tb, age_death,
    BYR, DYR, initstat, TB_yn
  ) |>
  filter(
    if_all(c(BYR), is.finite)
  ) |>
  mutate(
    TB_yn = as.numeric(as.character(TB_yn))
  ) |>
  drop_na(initstat)


match_full <- matchit(TB_yn ~ BYR + initstat,
                      data = match_full,
                      distance = "glm",
                      link = "probit")

matched_df <- match.data(match_full)

# Add constant status for lack of censoring
matched_df$status <- 1

# Create index date for non-TB
matched_df <- matched_df |>
  group_by(subclass) |>
  mutate(
    index_date = max(tb_first_date, na.rm = T)
  ) |>
  ungroup() |>
  mutate(
    index_date = if_else(is.infinite(index_date), NA_real_, index_date)
  ) |>
  drop_na(index_date) |>
  mutate(
    survival_time = DYR - index_date
  ) |>
  drop_na(survival_time) |>
  filter(survival_time >= 0)
  
# Final formatting of the matched data
plot_df <- matched_df |>
  mutate(
    event = 1
  ) |>
  mutate(TB_yn = factor(TB_yn, 
                        levels = c(0, 1), 
                        labels = c("Non-TB", "TB")))

# Fit km model
km_matched <- survfit(Surv(survival_time, event) ~ TB_yn, data = plot_df)

matched_p_final <- ggsurvplot(
  km_matched, 
  data = plot_df,
  conf.int = TRUE,       
  risk.table = FALSE,    
  pval = TRUE,  
  palette = c("#E7B800", "#2E9FDF"), 
  xlab = "Years of Life Remaining after Index Date",
  ggtheme = theme_minimal()
)

matched_p_final$plot <- matched_p_final$plot + 
  labs(title = "Mortality: TB vs. Matched Controls")

matched_p_final


# And table for the median survival times

medians <- surv_median(km_matched)

median_table <- medians |>
  mutate(
    Group = gsub("TB_tn=", "", strata),
    Median_Years = round(median, 2),
    Lower_95_CI = round(lower, 1),
    Upper_95_CI = round(upper, 1)
  ) |>
  select(Group, Median_Years, Lower_95_CI, Upper_95_CI)

print(median_table)


```

Now, to add more characteristics to matching:

-   Age at enlistment

-   War injuries

-   First regiment

-   Height at enlistment

-   Weight at enlistment

-   Occupation

```{r matching 2}


match_full <- surv_tb |>
  select(
    recidnum,
    tb_first_date, age_tb, age_death,
    BYR, DYR, initstat, initreg, gen_birthplace_icpsr1, wnd_yn,
    TB_yn
  ) |>
  filter(
    if_all(c(BYR), is.finite)
  ) |>
  mutate(
    TB_yn = as.numeric(as.character(TB_yn))
  ) |>
  drop_na(initstat, initreg, DYR, gen_birthplace_icpsr1)


match_full <- matchit(TB_yn ~ BYR + initstat + initreg + wnd_yn,
                      data = match_full,
                      method = "nearest",
                      distance = "glm",
                      link = "probit",
                      exact = ~initreg)


summary(match_full)

matched_df <- match.data(match_full)

# Add constant status for lack of censoring
matched_df$status <- 1

# Create index date for non-TB
matched_df <- matched_df |>
  group_by(subclass) |>
  mutate(
    index_date = if_else(
      any(!is.na(tb_first_date)), 
      max(tb_first_date, na.rm = TRUE), 
      NA_real_
    )
  ) |>
  ungroup() |>
  drop_na(index_date) |>
  mutate(
    survival_time = DYR - index_date
  ) |>
  drop_na(survival_time) |>
  filter(survival_time >= 0)
  
# Final formatting of the matched data
plot_df <- matched_df |>
  mutate(
    event = 1
  ) |>
  mutate(TB_yn = factor(TB_yn, 
                        levels = c(0, 1), 
                        labels = c("Non-TB", "TB")))

# Fit km model
km_matched <- survfit(Surv(survival_time, event) ~ TB_yn, data = plot_df)

matched_p_final <- ggsurvplot(
  km_matched, 
  data = plot_df,
  conf.int = TRUE,       
  risk.table = FALSE,    
  pval = TRUE,  
  palette = c("#E7B800", "#2E9FDF"), 
  xlab = "Years of Life Remaining after Index Date",
  ggtheme = theme_minimal()
)

matched_p_final$plot <- matched_p_final$plot + 
  labs(title = "Mortality: TB vs. Matched Controls")

matched_p_final


# And table for the median survival times

medians <- surv_median(km_matched)

median_table <- medians |>
  mutate(
    Group = gsub("TB_tn=", "", strata),
    Median_Years = round(median, 2),
    Lower_95_CI = round(lower, 1),
    Upper_95_CI = round(upper, 1)
  ) |>
  select(Group, Median_Years, Lower_95_CI, Upper_95_CI)

print(median_table)


# Cox model

cox_fit <- coxph(Surv(survival_time, event) ~ TB_yn, data = plot_df)

summary(cox_fit)


```
