---
title: "03_matching"
output: html_document
date: "2026-03-13"
---

# Matching

Based on the Kaplan-Meier curves made in the survival analysis markdown, we have decided to move forward with matching as a way of controlling for confounding variables. Throughout this markdown, the `survival` and `survminer` packages will be used.



```{r setup, include=FALSE}

library(tidyverse)
library(survival)
library(survminer)
library(data.table)
library(MatchIt)
library(knitr)
library(rlemon)

ntb <- fread("../data/in-progress/ntb_v6.csv")
tb <- fread("../data/in-progress/tb_v7.csv")
all <- rbind(tb, ntb)
surv_tb <- read_rds("../data/in-progress/surv_tb.RDS")
lst_df <- read_rds("../data/in-progress/lst_df.RDS")


############# Important functions ##############3#

remove_blank <- function(x){
  x <- x |>
    mutate(across(everything(), ~if_else(.x == "", NA, .x)))
  return(x)
}


# Index function (for after matching iteration and before graphing/model evaluation)
index <- function(x){
  # Turn matchit into dataframe 
  x <- match.data(x)
  
  # Constant status for lack of censoring
  x$status <- 1
  
  # Indexing
  x <- x |>
  group_by(subclass) |>
  mutate(
    index_date = max(tb_first_date, na.rm = T)
  ) |>
  ungroup() |>
  mutate(
    index_date = if_else(is.infinite(index_date), NA_real_, index_date)
  ) |>
  drop_na(index_date) |>
  mutate(
    survival_time = DYR - index_date
  ) |>
  drop_na(survival_time) |>
  filter(survival_time >= 0)
  
  x <- remove_blank(x)
  
  return(x)
}


# Function for manual checking (random sample of matches)
check_random_pairs <- function(df, n_pairs = 5){
  all_subclasses <- unique(df$subclass)
  n_pull <- min(n_pairs, length(all_subclasses))
  random_picks <- sample(all_subclasses, size = n_pull, replace = FALSE)
  df <- df |>
    filter(subclass %in% random_picks) |>
    arrange(subclass)
  
  return(df)
}

```

Matching will be used in this study due to the nature of our data and our research question. As mentioned in '02_survival-analysis.Rmd,' we are using KM curves not to evaluate survival after treatment, but to compare survival rates between soldiers with tuberculosis versus without. Therefore, our first event cannot be first mention of TB for both cohorts, since the non-TB group won't have a diagnosis date. We will use an `index_date` after matching the TB-afflicted soldiers to a healthy soldier with the most common traits.


## Matching Methodology

In order to have robust matching results, we have decided to go variable-by-variable for characteristics to match by. The first one will be birth year, to control for age. We will try to match by exact year and also by nearest neighbor.

```{r byr}

# Clean matching dataframe
match_byr <- surv_tb |>
  select(
    recidnum,
    tb_first_date, age_tb, age_death,
    BYR, DYR, initstat, TB_yn
  ) |>
  filter(
    if_all(c(BYR), is.finite)
  ) |>
  mutate(
    TB_yn = as.numeric(as.character(TB_yn)),
    BYR = as.numeric(as.character(BYR))
  ) |>
  drop_na(initstat, DYR, BYR) |>
  filter(!(TB_yn == 1 & is.na(tb_first_date)))


# Diagnostics before matching
nrow(match_byr |> filter(TB_yn == 1))
ggplot(match_byr, aes(x = BYR, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  labs(title = "Overlapping Distributions of BYR Between Cohorts")

# Run model
model_byr <- matchit(TB_yn ~ BYR,
                      data = match_byr,
                      distance = "glm",
                      link = "probit",
                      exact = ~BYR)



# INDEX
matched_df <- index(model_byr)

check_random_pairs(matched_df, n_pairs = 5)

summary((model_byr))


```

As we can see, there are n=1283 in the TB-cohort after cleaning NA dates, but before doing the matching. We will try to stay as close as possible to this number when matching as possible. The distributions are almost completely overlapping so this matching should be straightforward.

After matching, we have the same amount of pairs.

To minimize the number of people being thrown out with a 1:1 matching, we want to see whether a larger control group in the matching ratio is possible. This graph finds the sweet spot:

```{r byr_ratio}

# Define caliper
byr_caliper <- 1

# Define the maximum ratio
max_ratio <- 20
ratios_to_test <- 1:max_ratio

# Create an empty vector to store results
standardized_counts <- numeric(length(ratios_to_test))

# loop for each ratio
for (k in ratios_to_test) {

  m.out <- matchit(TB_yn ~ BYR, 
                   data = match_byr, 
                   method = "nearest", 
                   distance = "mahalanobis",
                   ratio = k,
                   caliper = c(BYR = 1),
                   std.caliper = FALSE)
  
  # Extract the number of matched control units
  total_matched_controls <- summary(m.out)$nn["Matched", "Control"]
  
  # Standardize the count
  standardized_counts[k] <- total_matched_controls / k
}

# Clean dataframe for
plot_data <- data.frame(
  Ratio = ratios_to_test,
  Standardized_Controls = standardized_counts
)

# Plot results
ggplot(data = plot_data, aes(x = Ratio, y = Standardized_Controls)) +
  geom_line(color = "#007acc", linewidth = 1.2) +
  geom_point(color = "#007acc", size = 4) +
  theme_minimal() +
  scale_x_continuous(breaks = ratios_to_test) +
  labs(
    title = "Standardized Matched Controls vs. Matching Ratio",
    subtitle = "Visualizing the drop-off of valid controls due to caliper limits",
    x = "Matching Ratio (1:k)",
    y = "Standardized Matched Controls (Total / k)"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold")
  )

```



```{r byr_final}

# We will use adjust this based on the matching results
byr_caliper <- 2

model_byr_caliper <- matchit(TB_yn ~ BYR,
                      data = match_byr,
                      distance = "glm",
                      link = "probit",
                      caliper = byr_caliper)


matched_df <- index(model_byr_caliper)
check_random_pairs(matched_df, n_pairs = 5)

summary(model_byr_caliper)

```

### Age at First Enlistment

Next, age at first enlistment:

```{r age_lst}


match_lst <- cbind(match_byr, "first_lst" = lst_df$first_lst, "age_lst" = lst_df$age_lst)
match_lst <- remove_blank(match_lst) |>
  drop_na(age_lst, initstat)


# lst diagnostics
nrow(match_lst |> filter(TB_yn == 1))
ggplot(match_lst, aes(x = age_lst, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  labs(title = "Overlapping Distributions of Age at First Enlistment Between Cohorts")


# Run matching algorithm
model_lst <- matchit(TB_yn ~ BYR + age_lst,
                      data = match_lst,
                      distance = "glm",
                      link = "probit",
                     caliper = c(age_lst = 2, BYR = 2),
                     ratio = 5)

# INDEX
matched_df <- index(model_lst)

check_random_pairs(matched_df, n_pairs = 5)
summary(model_lst)


```

And finally, height at enlistment:

```{r height}



```


Then state of first enlistment:

```{r initstat}

```



