---
title: "Matching Methodology"
author: "Giancarlo D'Agostino"
output:
  html_document:
    code_folding: hide
date: "2026-03-22"
---

# Matching

Based on the Kaplan-Meier curves made in the survival analysis markdown, we have decided to move forward with matching as a way of controlling for confounding variables. Throughout this markdown, the `survival` and `survminer` packages will be used.



```{r setup, include=FALSE}

library(tidyverse)
library(survival)
library(survminer)
library(data.table)
library(MatchIt)
library(knitr)
library(rlemon)
library(scales)

ntb <- fread("../data/in-progress/ntb_v6.csv")
tb <- fread("../data/in-progress/tb_v7.csv")
all <- rbind(tb, ntb)
surv_tb <- read_rds("../data/in-progress/surv_tb.RDS")
lst_df <- read_rds("../data/in-progress/lst_df.RDS")


############# Important functions ##############3#

remove_blank <- function(x){
  x <- x |>
    mutate(across(everything(), ~if_else(.x == "", NA, .x)))
  return(x)
}


# Index function (for after matching iteration and before graphing/model evaluation)
index <- function(x){
  # Turn matchit into dataframe 
  x <- match.data(x)
  
  # Constant status for lack of censoring
  x$status <- 1
  
  # Indexing
  x <- x |>
  group_by(subclass) |>
  mutate(
    tb_year = as.numeric(substr(as.character(tb_first_date), 1, 4)),
    index_year = suppressWarnings(max(tb_year, na.rm = TRUE))
  ) |>
  ungroup() |>
    
  mutate(
    index_year = if_else(is.infinite(index_year), NA_real_, index_year),
    DYR = as.numeric(as.character(DYR)),
    survival_time = DYR - index_year
  ) |>
  drop_na(index_year, survival_time) |>
  group_by(subclass) |>
  filter(survival_time >= 0) |>
  ungroup()
  
  x <- remove_blank(x)
  
  return(x)
}


# Function for manual checking (random sample of matches)
check_random_pairs <- function(df, n_pairs = 5){
  all_subclasses <- unique(df$subclass)
  n_pull <- min(n_pairs, length(all_subclasses))
  random_picks <- sample(all_subclasses, size = n_pull, replace = FALSE)
  df <- df |>
    filter(subclass %in% random_picks) |>
    arrange(subclass)
  
  return(df)
}


# Check matching ratio

check_matching_ratio <- function(match_formula, match_data, max_ratio = 10, my_calipers = NULL, my_exact = NULL) {
  
  ratios_to_test <- 1:max_ratio
  standardized_counts <- numeric(length(ratios_to_test))
  
  for (k in ratios_to_test) {
    # Run the matchit model dynamically based on the inputs provided
    m.out <- matchit(match_formula, 
                     data = match_data, 
                     method = "nearest", 
                     distance = "mahalanobis",
                     ratio = k,
                     caliper = my_calipers,
                     std.caliper = FALSE,
                     exact = my_exact)
    
    # Extract the number of matched control units
    total_matched_controls <- summary(m.out)$nn["Matched", "Control"]
    
    # Standardize the count
    standardized_counts[k] <- total_matched_controls / k
  }
  
  # Clean dataframe for plotting
  plot_data <- data.frame(
    Ratio = ratios_to_test,
    Standardized_Controls = standardized_counts
  )
  
  # Plot results
  p <- ggplot(data = plot_data, aes(x = Ratio, y = Standardized_Controls)) +
    geom_line(color = "#007acc", linewidth = 1.2) +
    geom_point(color = "#007acc", size = 4) +
    theme_minimal() +
    scale_x_continuous(breaks = ratios_to_test) +
    labs(
      title = "Standardized Matched Controls vs. Matching Ratio",
      # Dynamically update the subtitle so you know which variables you are testing!
      subtitle = paste("Formula:", deparse(match_formula)),
      x = "Matching Ratio (1:k)",
      y = "Standardized Matched Controls (Total / k)"
    ) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(face = "italic", color = "darkgray", size = 10),
      axis.title = element_text(face = "bold")
    )
  
  return(p)
}



```

Matching will be used in this study due to the nature of our data and our research question. As mentioned in '02_survival-analysis.Rmd,' we are using KM curves not to evaluate survival after treatment, but to compare survival rates between soldiers with tuberculosis versus without. Therefore, our first event cannot be first mention of TB for both cohorts, since the non-TB group won't have a diagnosis date. We will use an `index_date` after matching the TB-afflicted soldiers to a healthy soldier with the most common traits.


## Matching Methodology

In order to have robust matching results, we have decided to go variable-by-variable for characteristics to match by. The first one will be birth year, to control for age. We will try to match by exact year and also by nearest neighbor.

```{r byr}

# Clean matching dataframe
match_byr <- surv_tb |>
  select(
    recidnum,
    tb_first_date, age_tb, age_death,
    BYR, DYR, initstat, TB_yn, wnd_yn, lst1_hgt
  ) |>
  mutate(
    TB_yn = as.numeric(as.character(TB_yn)),
    BYR = as.numeric(as.character(BYR))
  ) |>
  drop_na(BYR, DYR) |> 
  filter(is.finite(BYR))


# Diagnostics before matching
nrow(match_byr |> filter(TB_yn == 1))
ggplot(match_byr, aes(x = BYR, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  labs(title = "Overlapping Distributions of BYR Between Cohorts")

# Run model
model_byr <- matchit(TB_yn ~ BYR,
                      data = match_byr,
                      distance = "glm",
                      link = "probit",
                      exact = ~BYR)



# INDEX
matched_df <- index(model_byr)

check_random_pairs(matched_df, n_pairs = 5)

summary((model_byr))


```

As we can see, there are n=1283 in the TB-cohort after cleaning NA dates, but before doing the matching. We will try to stay as close as possible to this number when matching as possible. The distributions are almost completely overlapping so this matching should be straightforward.

After matching, we have the same amount of pairs.

To minimize the number of people being thrown out with a 1:1 matching, we want to see whether a larger control group in the matching ratio is possible. This graph finds the sweet spot:

```{r byr_ratio}



check_matching_ratio(
  match_formula = TB_yn ~ BYR, 
  match_data = match_byr, 
  my_calipers = c(BYR = 1)
)


```



```{r byr_final}

# We will use adjust this based on the matching results
byr_caliper <- 2

model_byr_caliper <- matchit(TB_yn ~ BYR,
                      data = match_byr,
                      distance = "glm",
                      link = "probit",
                      caliper = byr_caliper)


matched_df <- index(model_byr_caliper)
check_random_pairs(matched_df, n_pairs = 5)

summary(model_byr_caliper)

```

### Age at First Enlistment

Next, age at first enlistment:

```{r age_lst}


match_lst <- cbind(match_byr, "first_lst" = lst_df$first_lst, "age_lst" = lst_df$age_lst)
match_lst <- remove_blank(match_lst) |>
  drop_na(age_lst)


# lst diagnostics
nrow(match_lst |> filter(TB_yn == 1))
ggplot(match_lst, aes(x = age_lst, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  labs(title = "Overlapping Distributions of Age at First Enlistment Between Cohorts")


# Run matching algorithm
model_lst <- matchit(TB_yn ~ BYR + age_lst,
                      data = match_lst,
                      distance = "glm",
                      link = "probit",
                     caliper = c(age_lst = 1, BYR = 1),
                     ratio = 5)

# INDEX
matched_df <- index(model_lst)

check_random_pairs(matched_df, n_pairs = 5)
summary(model_lst)

all$a_hgtft
```

```{r age_lst-ratio}

check_matching_ratio(
  match_formula = TB_yn ~ BYR + age_lst, 
  match_data = match_lst, 
  my_calipers = c(BYR = 1, age_lst = 1)
)

```


And finally, height at enlistment:

```{r height}

match_hgt <- remove_blank(match_lst) |>
  drop_na(lst1_hgt)

nrow(match_hgt |> filter(TB_yn == 1))
ggplot(match_hgt, aes(x = lst1_hgt, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  labs(title = "Overlapping Distributions of Height at First Enlistment Between Cohorts")


# Run matching algorithm
model_hgt <- matchit(TB_yn ~ BYR + age_lst + lst1_hgt,
                      data = match_hgt,
                      distance = "glm",
                      link = "probit",
                     caliper = c(age_lst = 1, BYR = 1),
                     ratio = 8
                    # exact = ~BYR + age_lst + lst1_hgt)
)
# INDEX
matched_df <- index(model_hgt)

check_random_pairs(matched_df, n_pairs = 5)
summary(model_hgt)

check <- as.data.frame(table(matched_df$subclass))
hist(check$Freq)


```

```{r height_ratio}

check_matching_ratio(
  match_formula = TB_yn ~ BYR + age_lst + lst1_hgt, 
  match_data = match_hgt, 
  my_calipers = c(BYR = 1, age_lst = 1, lst1_hgt = 1)
)

```



Then state of first enlistment:

```{r initstat}

match_state <- match_hgt
  

ggplot(match_state, aes(x = initstat, fill = factor(TB_yn))) + 
  geom_bar(position = "fill", color = "white", linewidth = 0.2) + 
  # Convert the y-axis
  scale_y_continuous(labels = percent_format()) + 
  scale_fill_manual(
    values = c("0" = "#4E79A7", "1" = "#E15759"), 
    labels = c("0" = "Non-TB", "1" = "TB")
  ) +
  labs(
    title = "Proportion of TB vs. Non-TB Recruits by Initial State",
    x = "Initial State",
    y = "Percentage within State",
    fill = "Cohort"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))


# Run matching algorithm
model_state <- matchit(TB_yn ~ BYR + age_lst + lst1_hgt + initstat,
                      data = match_hgt,
                      distance = "glm",
                      link = "probit",
                      caliper = c(age_lst = 1, BYR = 1, lst1_hgt = 1),
                      ratio = 8,
                      exact = ~initstat)

# INDEX
matched_df <- index(model_state)

check_random_pairs(matched_df, n_pairs = 20)
summary(model_state)

check <- as.data.frame(table(matched_df$subclass))
hist(check$Freq)


```

```{r initstat_ratio}

check_matching_ratio(
  match_formula = TB_yn ~ BYR + age_lst + lst1_hgt, 
  match_data = match_state, 
  my_calipers = c(BYR = 2, age_lst = 2, lst1_hgt = 2)
)

```


