---
title: "Matching Methodology"
author: "Giancarlo D'Agostino"
output:
  html_document:
    code_folding: hide
date: "2026-03-27"
---

# Matching

Based on the Kaplan-Meier curves made in the survival analysis markdown, we have decided to move forward with matching as a way of controlling for confounding variables. Throughout this markdown, the `MatchIt`, `survival`, and `survminer` packages will be used.



```{r setup, include=FALSE}

library(tidyverse)
library(survival)
library(survminer)
library(data.table)
library(MatchIt)
library(knitr)
library(rlemon)
library(scales)
library(patchwork)

ntb <- fread("../data/in-progress/ntb_v8.csv")
tb <- fread("../data/in-progress/tb_9.csv")
all <- rbind(tb, ntb)
surv_tb <- read_rds("../data/in-progress/surv_tb.RDS")
#lst_df <- read_rds("../data/in-progress/lst_df.RDS")


############# Important functions ###############

remove_blank <- function(x){
  if (is.data.frame(x)) {
    x <- x |>
    mutate(across(everything(), ~if_else(.x %in% c("", "."), NA, .x)))
  } else {
    x[x %in% c("", ".")] <- NA
  }
  return(x)
}


# Index function (for after matching iteration and before graphing/model evaluation)
index <- function(x){
  # Turn matchit into dataframe 
  x <- match.data(x)
  
  # Constant status for lack of censoring
  x$status <- 1
  
  # Indexing
  x <- x |>
  group_by(subclass) |>
  mutate(
    tb_year = as.numeric(substr(as.character(tb_first_date), 1, 4)),
    index_year = suppressWarnings(max(tb_year, na.rm = TRUE))
  ) |>
  ungroup() |>
    
  mutate(
    index_year = if_else(is.infinite(index_year), NA_real_, index_year),
    DYR = as.numeric(as.character(DYR)),
    survival_time = DYR - index_year
  ) |>
  drop_na(index_year, survival_time) |>
  group_by(subclass) |>
  filter(survival_time >= 0) |>
  ungroup()
  
  x <- remove_blank(x)
  
  return(x)
}


# Function for manual checking (random sample of matches)
check_random_pairs <- function(df, n_pairs = 5){
  all_subclasses <- unique(df$subclass)
  n_pull <- min(n_pairs, length(all_subclasses))
  random_picks <- sample(all_subclasses, size = n_pull, replace = FALSE)
  df <- df |>
    filter(subclass %in% random_picks) |>
    arrange(subclass)
  
  return(df)
}


# Check matching ratio

check_matching_ratio <- function(match_formula, match_data, max_ratio = 10, my_calipers = NULL, my_exact = NULL) {
  
  ratios_to_test <- 1:max_ratio
  standardized_counts <- numeric(length(ratios_to_test))
  
  for (k in ratios_to_test) {
    # Run the matchit model dynamically based on the inputs provided
    m.out <- matchit(match_formula, 
                     data = match_data, 
                     method = "nearest", 
                     distance = "mahalanobis",
                     ratio = k,
                     caliper = my_calipers,
                     std.caliper = FALSE,
                     exact = my_exact)
    
    # Extract the number of matched control units
    total_matched_controls <- summary(m.out)$nn["Matched", "Control"]
    
    # Standardize the count
    standardized_counts[k] <- total_matched_controls / k
  }
  
  # Clean dataframe for plotting
  plot_data <- data.frame(
    Ratio = ratios_to_test,
    Standardized_Controls = standardized_counts
  )
  
  # Plot results
  p <- ggplot(data = plot_data, aes(x = Ratio, y = Standardized_Controls)) +
    geom_line(color = "#007acc", linewidth = 1.2) +
    geom_point(color = "#007acc", size = 4) +
    coord_cartesian(ylim = c(0, 2200)) +
    theme_minimal() +
    scale_x_continuous(breaks = ratios_to_test) +
    labs(
      title = "Standardized Matched Controls vs. Matching Ratio",
      # Dynamically update the subtitle so you know which variables you are testing!
      subtitle = paste("Formula:", deparse(match_formula)),
      x = "Matching Ratio (1:k)",
      y = "Standardized Matched Controls (Total / k)"
    ) +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(face = "italic", color = "darkgray", size = 10),
      axis.title = element_text(face = "bold")
    )
  
  return(p)
}



```

Matching will be used in this study due to the nature of our data and our research question. As mentioned in '02_survival-analysis.Rmd,' we are using KM curves not to evaluate survival after treatment, but to compare survival rates between soldiers with tuberculosis versus without. Therefore, our first event cannot be first mention of TB for both cohorts, since the non-TB group won't have a diagnosis date. We will use an `index_date` after matching the TB-afflicted soldiers to a healthy soldier with the most common traits.


## Matching Methodology

In order to have robust matching results, we have decided to go variable-by-variable for characteristics to match by. The first one will be birth year, to control for age. We will try to match by exact year and also by nearest neighbor.

```{r byr}

# Clean matching dataframe
match_byr <- surv_tb |>
  select(
    recidnum,
    tb_first_date, age_tb, age_death,
    BYR, DYR, initstat, TB_yn, wnd_yn, lst1_hgt,
    lst1_jcd,
    bpl_cd = bpl_match_clean
  ) |>
  mutate(
    TB_yn = as.numeric(as.character(TB_yn)),
    BYR = as.numeric(as.character(BYR))
  ) |>
  drop_na(BYR, DYR) |> 
  filter(is.finite(BYR))


# Diagnostics before matching
nrow(match_byr |> filter(TB_yn == 1))
ggplot(match_byr, aes(x = BYR, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  labs(title = "Overlapping Distributions of BYR Between Cohorts")

# Run model
model_byr <- matchit(TB_yn ~ BYR,
                      data = match_byr,
                      distance = "glm",
                      link = "probit",
                      exact = ~BYR)



# INDEX
matched_df <- index(model_byr)

check_random_pairs(matched_df, n_pairs = 5)

summary((model_byr))


```

As we can see, there are n=1283 in the TB-cohort after cleaning NA dates, but before doing the matching. We will try to stay as close as possible to this number when matching as possible. The distributions are almost completely overlapping so this matching should be straightforward.

After matching, we have the same amount of pairs.

To minimize the number of people being thrown out with a 1:1 matching, we want to see whether a larger control group in the matching ratio is possible. This graph finds the sweet spot:

```{r byr_ratio, eval=FALSE}



check_matching_ratio(
  match_formula = TB_yn ~ BYR, 
  match_data = match_byr, 
  my_calipers = c(BYR = 1)
)


```



```{r byr_final}

# We will use adjust this based on the matching results
byr_caliper <- 1

model_byr_caliper <- matchit(TB_yn ~ BYR,
                      data = match_byr,
                      distance = "glm",
                      link = "probit",
                      caliper = byr_caliper)


matched_df <- index(model_byr_caliper)
check_random_pairs(matched_df, n_pairs = 5)

summary(model_byr_caliper)

```

### Age at First Enlistment

Next, age at first enlistment:

```{r age_lst}


match_lst <- cbind(match_byr, "first_lst" = all$first_lst, "age_lst" = all$age_lst)
match_lst <- remove_blank(match_lst) |>
  drop_na(age_lst)


# lst diagnostics
nrow(match_lst |> filter(TB_yn == 1))
ggplot(match_lst, aes(x = age_lst, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  labs(title = "Overlapping Distributions of Age at First Enlistment Between Cohorts")


# Run matching algorithm
model_lst <- matchit(TB_yn ~ BYR + age_lst,
                      data = match_lst,
                      distance = "glm",
                      link = "probit",
                     caliper = c(age_lst = 1, BYR = 1),
                     ratio = 5)

# INDEX
matched_df <- index(model_lst)

check_random_pairs(matched_df, n_pairs = 5)
summary(model_lst)


```

```{r age_lst-ratio, eval=FALSE}

check_matching_ratio(
  match_formula = TB_yn ~ BYR + age_lst, 
  match_data = match_lst, 
  my_calipers = c(BYR = 1, age_lst = 1)
)

```


And finally, height at enlistment (drop???):

```{r height}

match_hgt <- remove_blank(match_lst) |>
  drop_na(lst1_hgt)

nrow(match_hgt |> filter(TB_yn == 1))
ggplot(match_hgt, aes(x = lst1_hgt, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  labs(title = "Overlapping Distributions of Height at First Enlistment Between Cohorts")


# Run matching algorithm
model_hgt <- matchit(TB_yn ~ BYR + age_lst + lst1_hgt,
                      data = match_hgt,
                      distance = "glm",
                      link = "probit",
                     caliper = c(age_lst = 1, BYR = 1, lst1_hgt = 1),
                     ratio = 5
                    # exact = ~BYR + age_lst + lst1_hgt)
)
# INDEX
matched_df <- index(model_hgt)

check_random_pairs(matched_df, n_pairs = 5)
summary(model_hgt)

check <- as.data.frame(table(matched_df$subclass))
hist(check$Freq)


```

```{r height_ratio, eval=FALSE}

check_matching_ratio(
  match_formula = TB_yn ~ BYR + age_lst + lst1_hgt, 
  match_data = match_hgt, 
  my_calipers = c(BYR = 1, age_lst = 1, lst1_hgt = 1)
)

```



Then state of first enlistment:

```{r initstat}

match_state <- as.data.frame(match_hgt)
match_state$initstat <- remove_blank(match_state$initstat)

match_state <- match_state |>
  drop_na(initstat)
  

ggplot(match_state, aes(x = initstat, fill = factor(TB_yn))) + 
  geom_bar(position = "fill", color = "white", linewidth = 0.2) + 
  # Convert the y-axis
  scale_y_continuous(labels = percent_format()) + 
  scale_fill_manual(
    values = c("0" = "#4E79A7", "1" = "#E15759"), 
    labels = c("0" = "Non-TB", "1" = "TB")
  ) +
  labs(
    title = "Proportion of TB vs. Non-TB Recruits by Initial State",
    x = "Initial State",
    y = "Percentage within State",
    fill = "Cohort"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))


# Run matching algorithm
model_state <- matchit(TB_yn ~ BYR + age_lst + lst1_hgt + initstat,
                      data = match_state,
                      distance = "glm",
                      link = "probit",
                      caliper = c(age_lst = 1, BYR = 1, lst1_hgt = 1),
                      ratio = 8,
                      exact = ~initstat)

# INDEX
matched_df <- index(model_state)

check_random_pairs(matched_df, n_pairs = 20)
summary(model_state)

check <- as.data.frame(table(matched_df$subclass))
hist(check$Freq)


```

```{r initstat_ratio, eval=FALSE}

check_matching_ratio(
  match_formula = TB_yn ~ BYR + age_lst + lst1_hgt+ initstat, 
  match_data = match_state, 
  my_calipers = c(BYR = 1, age_lst = 1, lst1_hgt = 1)
)

```

```{r occupation}

match_occ <- match_state |>
  mutate(
    lst1_jcd = replace_na(as.character(lst1_jcd), "Missing"),
    lst1_jcd = as.factor(lst1_jcd)
  )

match_occ$lst1_jcd
  
model_occ <- matchit(TB_yn ~ BYR + age_lst + lst1_hgt + initstat + lst1_jcd,
                      data = match_occ,
                      distance = "glm",
                      link = "probit",
                      caliper = c(age_lst = 1, BYR = 1, lst1_hgt = 1),
                      ratio = 8,
                      exact = ~initstat + lst1_jcd)
  

matched_df <- index(model_occ)

check_random_pairs(matched_df, n_pairs = 20)
summary(model_occ)

```


```{r bpl}

match_bpl <- match_occ

match_bpl$bpl_cd
  
model_bpl <- matchit(TB_yn ~ BYR + age_lst + lst1_hgt + initstat + lst1_jcd + bpl_cd,
                      data = match_bpl,
                      distance = "glm",
                      link = "probit",
                      caliper = c(age_lst = 1, BYR = 1, lst1_hgt = 1),
                      ratio = 5,
                      exact = ~initstat + lst1_jcd)
  

matched_df <- index(model_bpl)

check_random_pairs(matched_df, n_pairs = 20)
summary(model_bpl)

check <- as.data.frame(table(matched_df$subclass))
hist(check$Freq)


```

We are losing too many people after birthplace, so we will remove it from the matching algorithm.

```{r match_final}

match_final <- match_occ

  
set.seed(2026)
model_final <- matchit(
  TB_yn ~ BYR + age_lst + lst1_hgt + initstat + lst1_jcd,
  data = match_final,
  distance = "glm",
  link = "logit",
  mahvars = ~ BYR + age_lst + lst1_hgt,   # enforce close numeric matching
  caliper = 0.2,                           # 0.2 * SD(logit(PS)) recommended default
  ratio = 5,
  #exact = ~ initstat + lst1_jcd,
  replace = FALSE,
  method = "nearest"                       # explicit; default is nearest
)
  

matched_df <- index(model_final)

check_random_pairs(matched_df, n_pairs = 5)
summary(model_final)

check <- as.data.frame(table(matched_df$subclass))
hist(check$Freq)

```







# Matching Diagnostics

Before fitting any models or analysis, we must first validate our matched cohorts using diagnostics.


```{r match_diagnostics}

library(cobalt)

love.plot(model_final, 
          treat = "TB_yn",
          binary = "std",       # Standardize the differences
          thresholds = 0.1,     # Draw the "golden rule" lines at 0.1 and -0.1
          drop.distance = TRUE, # Hides the propensity score row to focus on actual covariates
          var.order = "unadjusted", # Orders them by how unbalanced they were originally
          colors = c("red", "blue"),
          title = "Covariate Balance Before and After Matching")

bal.plot(model_final, 
         var.name = "distance", 
         which = "both", 
         type = "density", 
         colors = c("red", "blue"),
         title = "Propensity Score Distribution (Common Support)")

plot(model_final, type = "jitter", interactive = FALSE)
 
```



First we will check the balance between the TB and non-TB cohorts for each matching variable.

```{r cohort_comparison}



# 1. Extract the matched data from your MatchIt model
matched_data <- match.data(model_final)

# 2. Define standard colors/labels to keep code clean
cohort_colors <- c("0" = "#4E79A7", "1" = "#E15759")
cohort_labels <- c("0" = "Non-TB", "1" = "TB")

# ---------------------------------------------------------
# CONTINUOUS VARIABLE PLOTS (Density)
# ---------------------------------------------------------

p_byr <- ggplot(matched_data, aes(x = BYR, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = cohort_colors, labels = cohort_labels) +
  labs(title = "Birth Year (BYR)", x = "BYR", y = "Density", fill = "Cohort") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 10))

p_age <- ggplot(matched_data, aes(x = age_lst, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = cohort_colors, labels = cohort_labels) +
  labs(title = "Age at Enlistment", x = "Age", y = "Density", fill = "Cohort") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 10))

p_hgt <- ggplot(matched_data, aes(x = lst1_hgt, fill = factor(TB_yn))) + 
  geom_density(alpha = 0.5) +
  scale_fill_manual(values = cohort_colors, labels = cohort_labels) +
  labs(title = "Height", x = "Height", y = "Density", fill = "Cohort") +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", size = 10))


# ---------------------------------------------------------
# CATEGORICAL VARIABLE PLOTS (Stacked Percent)
# ---------------------------------------------------------

p_stat <- ggplot(matched_data, aes(x = initstat, fill = factor(TB_yn))) + 
  geom_bar(position = "fill", color = "white", linewidth = 0.2) + 
  scale_y_continuous(labels = percent_format()) + 
  scale_fill_manual(values = cohort_colors, labels = cohort_labels) +
  labs(title = "Initial State", x = "State", y = "Percentage", fill = "Cohort") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 6)
  )

p_jcd <- ggplot(matched_data, aes(x = lst1_jcd, fill = factor(TB_yn))) + 
  geom_bar(position = "fill", color = "white", linewidth = 0.2) + 
  scale_y_continuous(labels = percent_format()) + 
  scale_fill_manual(values = cohort_colors, labels = cohort_labels) +
  labs(title = "Job Code", x = "Job Code", y = "Percentage", fill = "Cohort") +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 6)
  )


# ---------------------------------------------------------
# COMBINE PLOTS WITH PATCHWORK
# ---------------------------------------------------------

# Collect the legend to the right so it isn't repeated 6 times
combined_plot <- (p_byr | p_age | p_hgt) / 
                 (p_stat | p_jcd) + 
  plot_layout(guides = "collect") + 
  plot_annotation(
    title = "Covariate Balance Diagnostics Post-Matching",
    theme = theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
                  plot.subtitle = element_text(size = 12, hjust = 0.5))
  )

# Render plot
print(combined_plot)

```


Finally, we will ask if our unmatched treated individuals and unmatched controls are significantly different than their matched counterparts.


```{r unmatched_comparison}

# 1. Extract matching weights to define "Matched" vs "Unmatched"
match_final_diag <- match_final %>%
  mutate(
    # If MatchIt assigned a weight > 0, they are in the matched sample
    matched_status = ifelse(model_final$weights > 0, "Matched", "Unmatched"),
    matched_status = factor(matched_status, levels = c("Unmatched", "Matched"))
  )

# 2. Split into Control (Non-TB) and Treated (TB) groups
data_control <- match_final_diag %>% filter(TB_yn == 0)
data_tb      <- match_final_diag %>% filter(TB_yn == 1)

# 3. Define variable lists
cont_vars <- c("BYR", "age_lst", "lst1_hgt")
cat_vars  <- c("initstat", "lst1_jcd")

# 4. Create plotting functions based on your specific aesthetic requests
plot_cont <- function(df, var_name, cohort_name) {
  ggplot(df, aes(x = .data[[var_name]], fill = matched_status)) +
    geom_density(alpha = 0.5) +
    # Using your preferred blue for matched, neutral gray for unmatched
    scale_fill_manual(values = c("Unmatched" = "gray70", "Matched" = "#4E79A7")) +
    labs(title = paste(var_name, "Distribution:", cohort_name), fill = "Status") +
    theme_minimal()
}

plot_cat <- function(df, var_name, cohort_name) {
  ggplot(df, aes(x = .data[[var_name]], fill = matched_status)) +
    geom_bar(position = "fill", color = "white", linewidth = 0.2) +
    scale_y_continuous(labels = percent_format()) +
    # Using your preferred red/blue theme
    scale_fill_manual(values = c("Unmatched" = "gray70", "Matched" = "#E15759")) +
    labs(
      title = paste("Proportion Matched by", var_name, "-", cohort_name),
      x = var_name,
      y = "Percentage within Category",
      fill = "Status"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
}

# 5. Generate the plots 
# (You can view them by typing control_cont_plots[[1]], etc., into the console)

# Control Group Plots (TB_yn == 0)
control_cont_plots <- lapply(cont_vars, function(v) plot_cont(data_control, v, "Control Group"))
control_cat_plots  <- lapply(cat_vars, function(v) plot_cat(data_control, v, "Control Group"))

# TB Group Plots (TB_yn == 1)
tb_cont_plots <- lapply(cont_vars, function(v) plot_cont(data_tb, v, "TB Group"))
tb_cat_plots  <- lapply(cat_vars, function(v) plot_cat(data_tb, v, "TB Group"))

control_cont_plots
control_cat_plots
tb_cont_plots
tb_cat_plots 

```


```{r download, include=FALSE}

matched_data <- index(model_final)

saveRDS(matched_data, file = "../data/in-progress/matched_data.RDS")

```





