---
title: "Survival Modeling"
author: "Giancarlo D'Agostino"
output: html_document
  html_document:
    code_folding: hide
date: "2026-05-03"
---

# Matching

After matching our cohorts, we are finally ready to model the survival of our Civil War soldiers. We will start by plotting Kalpan-Meier curves to see survival trajectory over time. After, we will fit a Cox regression with several covariates to see if geographic, age-related, demographic behavior, and/or other factors significantly decrease a person's chance of survival -- and if so, by how much.


```{r setup, include = FALSE}

library(tidyverse)
library(survival)
library(survminer)
library(data.table)
library(muhaz)

matched_data <- read_rds("../data/in-progress/matched_data.RDS")

ntb <- fread("../data/in-progress/ntb_v8.csv")
tb <- fread("../data/in-progress/tb_v9.csv")
all <- rbind(tb, ntb)

```



## Kaplan-Meier Curves


```{r km_curves}

matched_data <- matched_data |>
  mutate(event = 1)

# 2. Fit the Kaplan-Meier model
# We are modeling survival time by the TB_yn cohort
km_fit <- survfit(Surv(survival_time, event) ~ TB_yn, data = matched_data)

# 3. Plot the Kaplan-Meier curves using ggsurvplot
ggsurvplot(
  km_fit,
  data = matched_data,
  pval = TRUE,                     # Displays the p-value for the log-rank test
  conf.int = TRUE,                 # Displays 95% confidence intervals
  risk.table = TRUE,               # Adds a risk table below the plot
  risk.table.col = "strata",       # Colors the risk table numbers by cohort
  risk.table.height = 0.25,        # Adjusts how much space the table takes
  surv.median.line = "hv",         # Adds horizontal/vertical lines for median survival time
  palette = c("#4E79A7", "#E15759"), # Uses the blue (Non-TB) and red (TB) from your previous plots
  title = "Kaplan-Meier Survival Analysis: TB vs. Non-TB Cohort",
  xlab = "Years Since Index Date",
  ylab = "Survival Probability",
  legend.title = "Cohort",
  legend.labs = c("Non-TB", "TB"),
  ggtheme = theme_minimal()        # Keeps the aesthetic consistent
)

p1 <- ggsurvplot(
  km_fit, data = matched_data,
  xlim = c(0, 10),
  break.x.by = 1,                  # Forces risk table to show every year
  pval = TRUE, conf.int = TRUE, risk.table = TRUE,
  risk.table.col = "strata", risk.table.height = 0.25,
  surv.median.line = "hv",
  palette = c("#4E79A7", "#E15759"),
  censor.shape = 124,
  censor.size = 2,
  ggtheme = theme_minimal()
)

p1$plot <- p1$plot +
  scale_x_continuous(limits = c(0, 10), breaks = seq(0, 10, by = 1)) +
  scale_y_continuous(limits = c(0.7 , 1.0), breaks = seq(0.7, 1.0, by = 0.05), expand = c(0, 0)) +
  theme(panel.grid.minor = element_line(size = 0.25, colour = "grey90"))

print(p1)


p2 <- ggsurvplot(
  km_fit, data = matched_data,
  xlim = c(0, 5),
  break.x.by = 1,                  # Forces risk table to show every year
  pval = TRUE, conf.int = TRUE, risk.table = TRUE,
  risk.table.col = "strata", risk.table.height = 0.25,
  surv.median.line = "hv",
  palette = c("#4E79A7", "#E15759"),
  censor.shape = 124,
  censor.size = 2,
  ggtheme = theme_minimal()
)

p2$plot <- p2$plot +
  scale_x_continuous(limits = c(0, 5), breaks = seq(0, 5, by = 1)) +
  scale_y_continuous(limits = c(0.7 , 1.0), breaks = seq(0.7, 1.0, by = 0.05), expand = c(0, 0)) +
  theme(panel.grid.minor = element_line(size = 0.25, colour = "grey90"))

print(p2)

```

### Cumulative Hazard

```{r cum_haz}
# 5-Year Cumulative Hazard
ggsurvplot(
  km_fit,
  data = matched_data,
  fun = "cumhaz",
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE,
  risk.table.col = "strata",
  risk.table.height = 0.25,
  xlim = c(0, 5),
  ylim = c(0, 0.4),
  break.x.by = 1,                  # Risk table and x-axis breaks every year
  palette = c("#4E79A7", "#E15759"),
  title = "Cumulative Hazard (5-Year): TB vs. Non-TB Cohort",
  xlab = "Years Since Index Date",
  ylab = "Cumulative Hazard",
  legend.title = "Cohort",
  legend.labs = c("Non-TB", "TB"),
  ggtheme = theme_minimal()
)

# 10-Year Cumulative Hazard
ggsurvplot(
  km_fit,
  data = matched_data,
  fun = "cumhaz",
  pval = TRUE,
  conf.int = TRUE,
  risk.table = TRUE,
  risk.table.col = "strata",
  risk.table.height = 0.25,
  xlim = c(0, 10),
  ylim = c(0, 0.4),
  break.x.by = 1,                  # Risk table and x-axis breaks every year
  palette = c("#4E79A7", "#E15759"),
  title = "Cumulative Hazard (10-Year): TB vs. Non-TB Cohort",
  xlab = "Years Since Index Date",
  ylab = "Cumulative Hazard",
  legend.title = "Cohort",
  legend.labs = c("Non-TB", "TB"),
  ggtheme = theme_minimal()
)

```

```{r inst_haz}

# 5-Year Instantaneous Hazard Rate
tb_data <- matched_data[matched_data$TB_yn == 1, ]
ntb_data <- matched_data[matched_data$TB_yn == 0, ]

haz_tb_5yr <- muhaz(
  times = tb_data$survival_time,
  delta = tb_data$event,
  min.time = 0,
  max.time = 5
)

haz_ntb_5yr <- muhaz(
  times = ntb_data$survival_time,
  delta = ntb_data$event,
  min.time = 0,
  max.time = 5
)

plot(haz_tb_5yr, col = "#E15759", lwd = 2,
     main = "Instantaneous Hazard Rate (5-Year): TB vs. Non-TB",
     xlab = "Years Since Index Date",
     ylab = "Hazard Rate",
     xlim = c(0, 5))
lines(haz_ntb_5yr, col = "#4E79A7", lwd = 2)
legend("topright",
       c("TB", "Non-TB"),
       col = c("#E15759", "#4E79A7"),
       lwd = 2)


# 10-Year Instantaneous Hazard Rate
haz_tb_10yr <- muhaz(
  times = tb_data$survival_time,
  delta = tb_data$event,
  min.time = 0,
  max.time = 10
)

haz_ntb_10yr <- muhaz(
  times = ntb_data$survival_time,
  delta = ntb_data$event,
  min.time = 0,
  max.time = 10
)

plot(haz_tb_10yr, col = "#E15759", lwd = 2,
     main = "Instantaneous Hazard Rate (10-Year): TB vs. Non-TB",
     xlab = "Years Since Index Date",
     ylab = "Hazard Rate",
     xlim = c(0, 10))
lines(haz_ntb_10yr, col = "#4E79A7", lwd = 2)
legend("topright",
       c("TB", "Non-TB"),
       col = c("#E15759", "#4E79A7"),
       lwd = 2)

```


## Cox Regression

1 year survival, 5 year survival, 10 year (survival table / life table)

cumulative hazard curve

instananeous hazard

why is the slope steadying out?
- severe tb cases die immediately &
- less severe, fights back OR
- misclassified (pneumonia)

- people with low immune systems.
  - after first 5 years, who are the ones that die?
     - table 1: people that died (which factors are related to death?) vs. survived
     
- table one: description of the cohort

```{r cox}


# 1. Fit the basic Cox regression model
# We use the same Surv() object we created for the Kaplan-Meier curve
cox_model <- coxph(Surv(survival_time, event) ~ TB_yn, data = matched_data)

# 2. View the results
summary(cox_model)


```

# Cox For First 10 Years



